\\fontsize{12pt}{3pt}\\selectfont \\textbf{\\textit{Submitted to Ecosphere. Manuscript type: Article}}

### 2023-10-19 manuscript/manuscriptRW.Rmd
changes in manuscript before removing figures

      $^3$Department of Computational Hydrosystem, Helmholtz-Centre for Environmental Research - UFZ, Permoserstraße 15, 04318 Leipzig, Germany. \newline

@@

       $^{10}$Department of Forest Harvesting, Logistics and Ameliorations, Faculty of Forestry, Technical University in Zvolen \newline

@@

       $^{13}$Dpt. of Soil and Climate, Bavarian State Institute of Forestry. \newline 

@@

    number_sections: yes

@@

Most of the models overestimate recruitment levels, which is only compensated in some models by high simulated mortality rates in the early stages of individual trees dynamics. Simulated species diversity of recruitment matches the observed ranges. Models simulating higher species diversity at the stand level do not present higher recruitment diversity. The effect of light availability on recruitment levels is captured better than the effect of temperature and soil moisture, but patterns are not consistent across models. Increasing complexity in the tree regeneration modules of the models is not related to higher accuracy of simulated tree recruitment. Furthermore, individual model design is more important than scale (stand, landscape, global) and approach (empirical, process-based) for accurately capturing tree regeneration.

@@

Overall, models are needed to (1) synthesize existing empirical data and explore their relationships, (2) assess future tree regeneration e.g. in the context of global change scenarios, and (3) identify the most important processes that are shaping ecological patterns. Given the current strategies that are used in models of forest dynamics to represent tree regeneration, model behavior often is prone to problems, such as very high levels of tree regeneration that necessitate excess mortality at early stages of tree life to simulate correct stand structure and composition. Also, inadequately high species diversity in tree regeneration may be simulated, which is characteristic of "classical" forest gap models [@gutiérrez2016], at least as long as the simulation set-up comprises a multitude of species. Some models use calibration against local canopy-level data to constrain simulated regeneration levels, which is likely to hamper the general applicability of these models e.g. under scenarios of climate change. Furthermore, correctly capturing the species composition in tree regeneration as a function of the presence of seed trees in the canopy is often a particular challenge, potentially leading to unrealistic successional drift in the model, which must be corrected by factors that are hard to parameterize [e.g., @lischke2006]. A related issue is the excessive reduction of species diversity due to positive feedback effects, such that eventually just single-species stands remain. This is sometimes corrected by the incorporation of a low level of seed influx of all species at all times [@schumacher2006] or by restricting the number of seeds per species in the seedbank [@lischke2006]. However, simulated species composition is usually exceedingly sensitive to assumptions about seed availability, whereas the parameters of such functions are poorly constrained by field data. Lastly, there is often a problem with insufficient observational constraints on parameter values for models that start from very small tree sizes (e.g., 10 cm of height or even seeds) and track tree development in a process-oriented manner by considering a multitude of ecological influences, rather than emphasizing tree regeneration into a larger size class [e.g., @wehrli2006].

Modeling tree regeneration processes is challenging, and even in empirical ecology, it has not received much attention [@hanbury-brown2022], although many sub-processes have been studied in detail [@miina2006]. Yet, little data are available that covers all the processes within one species along environmental gradients, let alone for a vast suite of species. Data on forest regeneration data are often fragmented, which constitutes a major problem for model building [@clark1999a]. For example, monitoring on permanent plots (such as National Forest Inventories) often measures tree regeneration [@zell2019], however with a design that captures the rate after the stand initiation phase [@hallsby2015]. Data from permanent plots, although available across regions, are also highly heterogeneous, rendering their use difficult in a modeling context [@käber2021], and targeted studies to measure tree regeneration on experimental sites are often limited in spatial extent, species studied, or the sub-set of processes that are targeted [@berdanier2016; @collet2006].

@@

More specifically, we aim to answer the following questions: (1) Are models of forest dynamics capturing accurately tree recruitment levels, initial tree species diversity and mortality in the recruitment? (2) Do model traits explain differences in model performance? (3) How well do the models capture total recruitment and individual species regeneration niches across environmental gradients of light availability, temperature, and soil moisture? We evaluate the performance of regeneration models by analyzing simulated data alongside observed data. Specifically, we examine tree species diversity, mortality rates in tree recruitment, and recruitment levels. Then, we link the model's traits and performance to define regeneration and we assess the recruitment gradients on regeneration niches. Finally, we provide research recommendations for next steps towards improving tree regeneration modules.

@@

Recruitment data covering a wide range of environmental conditions are hard to obtain, and this is one of the reasons why most models of forest dynamics have never been confronted with a dataset covering such gradients over a large number of sites, to evaluate how well regeneration is captured. The observations used here are derived from a novel network of sites in forest reserves that represent the range of environmental gradients in temperature and precipitation in Central Europe as compiled in the framework of the EuFoRIa network [@euforia2019;@käber2023] (Figure \@ref(fig:map)). The environmental coverage and the variety of forests represented by this data is unprecedented within Europe. These forests have been unmanaged for at least 10 years prior to their designation as reserves, and most of them provide time series of natural forest dynamics over multiple decades and up to ca. 80 years. The census periods range from 3 to 37 years, with an average of 14 years. The data provide information at the tree level, thus allowing for the sequential comparison of processes such as individual-tree regeneration and death between the measurements.

We selected 200 sites from this network as the benchmarking dataset for the simulation to be representative of the environmental variation contained in the data. This was achieved by applying k-means clustering to define 200 clusters of plots from the original set of 869 plots along the environmental dimensions of temperature, climatic water balance, soil quality, slope and aspect. Each of these sites featured at least two consecutive measurements. Recruitment minimum size thresholds for these sites differed between stem diameters of 0 and 10 cm. For our study, we defined two datasets where one included 165 sites with a diameter threshold of 7 cm or lower, and the other included another 35 sites with diameter thresholds from 10 cm. Therefore we had in total 200 sites with observations of newly established trees above 10 cm and 165 sites above 7cm. In the results section, any result from observed data at 7 cm threshold is originated from the 165 sites.

@@

The simulations were run for mixed-species forests (not multiple single-species simulations) using mixtures of eleven species or genera for which regeneration data of sufficient quality were available from EuFoRIa: *Fagus sylvatica L., Picea abies L., Abies alba Mill., Carpinus betulus L., Tilia cordata Mill., Acer pseudoplatanus L., Betula spp. L., Fraxinus excelsior L., Quercus spp. L., Alnus glutinosa L., Pinus sylvestris L.* The same set of species was used at all 200 sites. Two models included their standard set of species for the simulations, which is much larger (i.e., ForClim 1, ForClim 11, TreeMig), the species simulated beyond the eleven species were lumped in a category "others". In three models, fewer than these eleven species were simulated (4C, xComp, LPJ-GUESS) (Table \@ref(tab:tableModels)). In 4C only *Fagus sylvatica*, *Picea abies*, *Betula* spp., *Quercus* spp. and *Pinus sylvestris* are parameterised. xComp simulations did not consider *Fraxinus excelsior* due to a depreciated species parameterisation. In LPJ-GUESS, *Acer pseudoplatanus* and *Alnus glutinosa* are not parameterized and therefore these species could not be included in the simulation. Finally, the model aDGVM2 does not simulate individual species. Rather, community assembly processes and trait filtering generate plant communities that are adjusted to the biotic and abiotic conditions, and the simulated plants can be classified into ecological strategies based on their trait values in a post-processing step.

@@

```{=tex}
\begin{equation}
H_{R_{n}} = - \sum_{i = 1}^{s}{p_{R_{n,i}}\ln(p_{R_{n,i}})}\label{eqpiShannonR}
\end{equation}
```
```{=tex}
\begin{equation}
p_{R_{n,i}} = \frac{r_{BA_{n,i}}}{TotalR_{BA_{n}}} \label{eqpiShannonR2}
\end{equation}
```
```{=tex}
\begin{equation}
H_{S_{n}} = - \sum_{i = 1}^{s}{p_{S_{n,i}}\ln(p_{S_{n,i}})}\label{eqpiShannon}
\end{equation}
```
```{=tex}
\begin{equation}
p_{S_{n,i}} = \frac{BA_{n,i}}{Total_{BA_{n}}} \label{eqpiShannon2}
\end{equation}
```
where $s$ is the total number of species present and that have a basal area larger than zero in sample $n$; $p_{R_{n,i}}$ is the proportion of species $i$ in sample $n$ calculated as the regeneration basal area ($r_{BA_{n,i}}$) for that species $i$ relative to total recruited basal area ($TotalR_{BA_{n}}$) of the sample $n$; $p_{S_{n,i}}$ is the proportion of species $i$ calculated as the basal area of all trees ($BA_{n,i}$) of that species $i$ relative to total basal area ($TotalR_{BA_{n}}$) of the sample $n$. Species diversity was not assessed for aDGVM2 as this model does not simulate individual species.

@@

Simulated levels of recruitment varied strongly across the 15 models and typically did not match the levels found in the forest reserve data (Figure \@ref(fig:over)). Recruitment was overestimated in most models for both the 7 and 10 cm diameter thresholds, with the exception of the empirical stand model SIBYLA, the landscape model Landis-II and the global model aDGVM2, which estimated recruited levels at the lower end of the plausibility interval of the observed data (Figure \@ref(fig:over)). The models with the largest overestimation were the stand model PICUS and the landscape model TreeMig. For most models, the variability of simulated regeneration levels across the 200 sites (visible from the interquartile range in the box plots of Figure \@ref(fig:over)) was similar to or smaller than observed, with the notable exception of PICUS, where simulated recruitment variability was much larger.

@@

There was no significant relationship between the mean complexity of the regeneration module (Table \@ref(tab:tableModels)) and the overestimation proportion at a diameter threshold of 7 cm (Figure S6). All the models except 4C had significant differences between the observed and simulated mean regeneration values. Therefore it was not possible to differentiate in terms of performance between models type (empirical or process based), or their scales (stand, landscape or global) (Table S5). The values of recruited species diversity simulated by each model were also assessed against the observed data to see if there was a difference between models that include a feedback compared to those that did not; all the models except ForCEEPs(f) and FORMIND had significant differences (Table S6). It is noteworthy that for the model ForCEEPS the pattern changed from the overprediction of species diversity in the regeneration to a diversity level that is closer to observations when the recruitment module included feedback (ForCEEPs(f)).

@@

When evaluating recruited levels along key gradients of light availability (basal area), temperature (degree-days) and soil moisture (climatic water balance), distinct features emerged: the models reproduced the effect of basal area in both its magnitude and patterns (Figure \@ref(fig:nvTrend710BAlightTemp) and S7) considerably better than the effects of the climatic gradients, where they featured varying patterns.

@@

The observed recruitment did not change much across the climatic water balance gradient, and showed a slight increase with temperature. The models did not match the observed reruitment patterns across the gradients of soil moisture and temperature (Figure \@ref(fig:nvTrend710BAlightTemp)b,c), and neither was there a clear pattern across models. Regarding the soil moisture gradient, a group of models featured decreasing recruitment with increasing soil moisture (4C, ForClim 1, ForClim11, SIBYLA, xComp, PICUS, LandClim, TreeMig and iLand), while a few models showed the opposite trend (ForCEEPS(f), FORMIND and LPJ-GUESS). Across the temperature gradient, all the models that featured a decrease in recruitment levels with increasing soil moisture showed the opposite trend (i.e., more recruitment with increasing temperature), except for 4C. The other models did not feature a clear pattern.

@@

The EuFoRIa data [@käber2023] are exceptional, particularly with respect to the number of records (number of sites and repeated measurements on unmanaged forests), which is essential for capturing a highly 'noisy' process such as tree regeneration. The use of this dataset for model benchmarking provided novel insights on the ability of state-of-the-art models to accurately simulate regeneration levels, species composition, and mortality in an early stage of tree life, i.e., between a tree diameter of 7 and 10 cm. Overall, by adopting this approach a much broader understanding resulted than if we had used these data for model calibration: it is primarily from the shortcomings of the models that we can gain ecological insights [cf. @trugman2022].

@@

Regeneration levels define the structure and composition of future forests, but it is equally important to correctly identify areas where regeneration is lacking [@rammer2021]. There are multiple constraints to the regeneration niche of tree species [@price2001], and therefore the absence of regeneration is likely to be common [@fortin2007], even over larger areas such as the one-hectare plots used here. Tree regeneration data are characterized by zero-inflation. This was clearly evident from the EuFoRIa dataset, but some of the models did not produce zeros at all, or featured a very low proportion of zero data. This substantial difference may be due to the fact that the simulation results were drawn from equilibrium forests, whereas in reality many of the forest reserves are recovering from past management activities and have become denser over the past decades [e.g., @heiri2009], leading to less regeneration than in an equilibrium situation. Another possible reason for this difference may be the exclusion of factors like deer browsing or failure to accurately measure establishment filters, such as those to site limitants or competition with herb layers.

@@

The species diversity of the entire stand was captured better by the models than the diversity of tree recruitment, and only a few models overpredicted stand-level diversity. Defining regeneration composition as being proportional to adult abundance, regardless of productivity, might be a simple and conservative assumption to safely maintain relative species abundances [@hanbury-brown2022], although this may be simplistic. Furthermore, based on the simulation results, there is no evidence that models with feedback from the canopy (in terms of species composition of regeneration) captured better the species diversity of the recruitment than those without feedback, with the exception of ForCEEPS that significantly improved regeneration species diversity with the model version that included feedback (ForCEEPS(f)). The similar performance of models with and without feedback is likely because the models put more weight on the regeneration niche arising from abiotic and biotic filters, than from the habitat niche of the adult trees [@grubb1977].

@@

There were pronounced differences how the main tree species were represented by the models along the environmental gradients, in particular the dominant species *Abies alba*, *Fagus sylvatica* and *Picea abies*. The regeneration levels were sampled from simulations in the equilibrium, and in this state it is expected that non-climax species such as *Pinus sylvestris* or *Quercus* would be of minor importance, or absent [@klopcic2015]. Most models captured this low abundance, which is also found in the empirical data. Thus, although the broad patterns are matched by many models, improvements in the quantification of the regeneration niche of the species are needed, but this cannot be done in the absence of robust datasets across multiple species.

@@

However, most models properly capture the diversity of the initial tree community, and differences between model formulations, e.g., the presence or absence of feedback from the adult trees, did not have a strong effect for capturing the species composition of regeneration.

@@

**Table 1.** Models included in this study, their characteristics and simulation strategies used. Scale refers to the models scale, Type to the models type, Population structure to the model population approach, Feedback to the inclusion or not of adult tree feedback to the level and species composition of regeneration, Approach to the model regeneration module approach, Start from to the stage from which trees are recruited, Complexity, to the mean regeneration formulation as estimated in (Bugmann & Seidl, 2022), Species to which species from the original eleven species were simulated, Runtime (spin up) [years] to the number of years used per model in their simulations for the spin up period and regeneration sampling; Climate data type (use of years) to the type of data used in the simulations and how it was used, and Reference provides the link to the publication with more model details. Further information can be found on the individual model reports available in the supplementary materials.

@@

**Figure 7.** Mean regeneration values across the 200 samples per site with the y-axis scaled differently by model, for the 200 sites against gradients of (a) total basal area; (b) climatic water balance; (c) seasonal degree-day sum. The values were split into ten bins; the red lines represent a GAM model showing the trend in the observed data. Y-axis are not equally scaled.

>>>>

      $^3$Department of Ecological Modelling, Helmholtz-Centre for Environmental Research - UFZ, Permoserstraße 15, 04318 Leipzig, Germany. \newline

@@

       $^{10}$Department of Forest Harvesting, Logistics and Ameliorations, Faculty of Forestry, Technical University in Zvolen, Slovakia \newline

@@

       $^{13}$Dpt. of Soil and Climate, Bavarian State Institute of Forestry, Germany. \newline 

@@

    number_sections: no

@@

Most of the models overestimate recruitment levels, which is only
compensated in some models by high simulated mortality rates in the
early stages of individual trees dynamics. Simulated species diversity
of recruitment matches the observed ranges. Models simulating higher
species diversity at the stand level do not present higher species
diversity in the recruitment layer. The effect of light availability on
recruitment levels is captured better than the effect of temperature and
soil moisture, but patterns are not consistent across models. Increasing
complexity in the tree regeneration modules of the models is not related
to higher accuracy of simulated tree recruitment. Furthermore,
individual model design is more important than scale (stand, landscape,
global) and approach (empirical, process-based) for accurately capturing
tree regeneration.

@@

Overall, models are needed to (1) synthesize existing empirical data and
explore their relationships, (2) assess future tree regeneration e.g. in
the context of global change scenarios, and (3) identify the most
important processes that are shaping ecological patterns. Given the
current strategies that are used in models of forest dynamics to
represent tree regeneration, model behavior often is prone to problems,
such as very high levels of tree regeneration that necessitate excess
mortality at early stages of tree life to simulate correct stand
structure and composition [@kroiss2015]. Also, inadequately high species
diversity in tree regeneration may be simulated, which is characteristic
of "classical" forest gap models [@gutiérrez2016], at least as long as
the simulation set-up comprises a multitude of species. Some models use
calibration against local canopy-level data to constrain simulated
regeneration levels, which is likely to hamper the general applicability
of these models e.g. under scenarios of climate change. Furthermore,
correctly capturing the species composition in tree regeneration as a
function of the presence of seed trees in the canopy is often a
particular challenge, potentially leading to unrealistic successional
drift in the model, which must be corrected by factors that are hard to
parameterize [e.g., @lischke2006]. A related issue is the excessive
reduction of species diversity due to positive feedback effects, such
that eventually just single-species stands remain [@meier2011]. This is
sometimes corrected by the incorporation of a low level of seed influx
of all species at all times [@schumacher2006] or by restricting the
number of seeds per species in the seedbank [@lischke2006]. However,
simulated species composition is usually exceedingly sensitive to
assumptions about seed availability, whereas the parameters of such
functions are poorly constrained by field data. Lastly, there is often a
problem with insufficient observational constraints on parameter values
for models that start from very small tree sizes (e.g., 10 cm in height
or even seeds) and track tree development in a process-oriented manner
by considering a multitude of ecological influences, rather than
emphasizing tree regeneration into a larger size class [e.g.,
@wehrli2006].

Modeling tree regeneration processes is challenging, and even in
empirical ecology, it has not received much attention
[@hanbury-brown2022], although many sub-processes have been studied in
detail [@miina2006]. Yet, little data are available that cover all the
processes within one species along environmental gradients, let alone
for a vast suite of species. Data on forest regeneration are often
fragmented, which constitutes a major problem for model building
[@clark1999a]. For example, monitoring on permanent plots (such as
National Forest Inventories) often measures tree regeneration
[@zell2019], however with a design that captures the rate after the
stand initiation phase [@hallsby2015]. Data from permanent plots,
although available across regions, are also highly heterogeneous,
rendering their use difficult in a modeling context [@käber2021], and
targeted studies to measure tree regeneration on experimental sites are
often limited in spatial extent, species studied, or the sub-set of
processes that are targeted [@berdanier2016; @collet2006].

@@

More specifically, we aim to answer the following questions: (1) Are
models of forest dynamics capturing accurately tree recruitment levels,
initial tree species diversity and mortality in the recruitment? (2) Do
model traits explain differences in model performance? (3) How well do
the models capture total recruitment and individual species regeneration
niches across environmental gradients of light availability,
temperature, and soil moisture? We evaluate the performance of
regeneration models by analyzing simulated data alongside observed data.
Specifically, we examine tree species diversity, mortality rates in tree
recruitment, and recruitment levels. Then, we link model traits (e.g.,
complexity, scale) to performance, and we assess whether modeled total
recruitment or individual species regeneration niches (i.e.,
proportional recruitment along an environmental gradient) align with
observational data.

@@

Recruitment data covering a wide range of environmental conditions are
hard to obtain, and this is one of the reasons why most models of forest
dynamics have never been confronted with a dataset covering such
gradients over a large number of sites, to evaluate how well
regeneration is captured. The observations used here are derived from a
novel network of sites in forest reserves that represent the range of
environmental gradients in temperature and precipitation in Central
Europe as compiled in the framework of the EuFoRIa network
[@euforia2019; @käber2023] (Figure \@ref(fig:map)). The environmental
coverage and the variety of forests represented by this data is
unprecedented within Europe. These forests have been unmanaged for at
least 10 years prior to their designation as reserves, and most of them
provide time series of natural forest dynamics over multiple decades and
up to ca. 80 years. The census periods range from 3 to 37 years, with an
average of 14 years. The data provide information at the tree level,
thus allowing for the sequential comparison of processes such as
individual-tree regeneration and death between the measurements.

We selected 200 sites from this network as the benchmarking dataset for
the simulation to be representative of the environmental variation
contained in the data. This was achieved by applying k-means clustering
to define 200 clusters of plots from the original set of 869 plots along
the environmental dimensions of temperature, climatic water balance,
soil quality, slope and aspect. Each of these sites featured at least
two consecutive measurements. Recruitment minimum size thresholds for
these sites differed between stem diameters of 0 and 10 cm. For our
study, we defined two datasets where one included 165 sites with a
diameter threshold of 7 cm or lower, and the other included another 35
sites with diameter thresholds from 10 cm. Therefore we had in total 200
sites with observations of newly established trees above 10 cm and 165
sites above 7cm. In the results section, any result from observed data
at 7 cm threshold originated from the 165 sites.

@@

The simulations were run for mixed-species forests (not multiple
single-species simulations) using mixtures of eleven species or genera
for which regeneration data of sufficient quality were available from
EuFoRIa: *Fagus sylvatica L., Picea abies L., Abies alba Mill., Carpinus
betulus L., Tilia cordata Mill., Acer pseudoplatanus L., Betula spp. L.,
Fraxinus excelsior L., Quercus spp. L., Alnus glutinosa L., Pinus
sylvestris L.* The same set of species was used at all 200 sites. Two
models included their standard set of species for the simulations, which
is much larger (i.e., ForClim 1, ForClim 11, TreeMig), the species
simulated beyond the eleven species were lumped in a category "others".
In three models, fewer than these eleven species were simulated (4C,
xComp, LPJ-GUESS) (Table \@ref(tab:tableModels)). In 4C only *Fagus
sylvatica*, *Picea abies*, *Betula* spp., *Quercus* spp. and *Pinus
sylvestris* are parameterised. xComp simulations did not consider
*Fraxinus excelsior* due to a deprecated species parameterisation. In
LPJ-GUESS, *Acer pseudoplatanus* and *Alnus glutinosa* are not
parameterized and therefore these species could not be included in the
simulation. Finally, the model aDGVM2 does not simulate individual
species. Rather, community assembly processes and trait filtering
generate plant communities that are adjusted to the biotic and abiotic
conditions, and the simulated plants can be classified into ecological
strategies based on their trait values in a post-processing step.

@@

```{=tex}
\begin{equation}
H_{R_{n}} = - \sum_{i = 1}^{s}{p_{R_{n,i}}\ln(p_{R_{n,i}})}\label{eqpiShannonR}
\end{equation}
```
```{=tex}
\begin{equation}
p_{R_{n,i}} = \frac{r_{BA_{n,i}}}{TotalR_{BA_{n}}} \label{eqpiShannonR2}
\end{equation}
```
```{=tex}
\begin{equation}
H_{S_{n}} = - \sum_{i = 1}^{s}{p_{S_{n,i}}\ln(p_{S_{n,i}})}\label{eqpiShannon}
\end{equation}
```
```{=tex}
\begin{equation}
p_{S_{n,i}} = \frac{BA_{n,i}}{Total_{BA_{n}}} \label{eqpiShannon2}
\end{equation}
```
where $s$ is the total number of species present that have a basal area
larger than zero in sample $n$; $p_{R_{n,i}}$ is the proportion of
species $i$ in sample $n$ calculated as the regeneration basal area
($r_{BA_{n,i}}$) for that species $i$ relative to total recruited basal
area ($TotalR_{BA_{n}}$) of the sample $n$; $p_{S_{n,i}}$ is the
proportion of species $i$ calculated as the basal area of all trees
($BA_{n,i}$) of that species $i$ relative to total basal area
($Total_{BA_{n}}$) of the sample $n$. Species diversity was not assessed
for aDGVM2 as this model does not simulate individual species.

@@

Simulated levels of recruitment varied strongly across the 15 models and
typically did not match the levels found in the forest reserve data
(Figure \@ref(fig:over)). Recruitment was overestimated in most models
for both the 7 and 10 cm diameter thresholds, with the exception of the
empirical stand model SIBYLA, the landscape model Landis-II and the
global model aDGVM2, which estimated recruited levels at the lower end
of the plausibility interval of the observed data (Figure
\@ref(fig:over)). The models with the largest overestimation were the
stand model PICUS and the landscape model TreeMig. For most models, the
variability of simulated recruitment levels across the 200 sites
(visible from the interquartile range in the box plots of Figure
\@ref(fig:over)) was similar to or smaller than observed, with the
notable exception of PICUS, where simulated recruitment variability was
much larger.

@@

There was no significant relationship between the mean complexity of the
regeneration module (Table \@ref(tab:tableModels)) and the
overestimation proportion at a diameter threshold of 7 cm (Figure S6).
All the models except 4C had significant differences between the
observed and simulated mean recruitment values. Therefore it was not
possible to differentiate in terms of performance between models type
(empirical or process based), or their scales (stand, landscape or
global) (Table S5). The values of recruited species diversity simulated
by each model were also assessed against the observed data to see if
there was a difference between models that include a feedback compared
to those that did not; all the models except ForCEEPs(f) and FORMIND had
significant differences (Table S6). It is noteworthy that for the model
ForCEEPS the pattern changed from the overprediction of species
diversity in the regeneration to a diversity level that is closer to
observations when the recruitment module included feedback
(ForCEEPs(f)).

@@

When evaluating total recruited levels along key gradients of light
availability (basal area), temperature (degree-days) and soil moisture
(climatic water balance), distinct features emerged: the models
reproduced the effect of basal area in both its magnitude and patterns
(Figure \@ref(fig:nvTrend710BAlightTemp) and S7) considerably better
than the effects of the climatic gradients, where they featured varying
patterns.

@@

The observed recruitment did not change much across the climatic water
balance gradient, and showed a slight increase with temperature. The
models did not match the observed recruitment patterns across the
gradients of soil moisture and temperature (Figure
\@ref(fig:nvTrend710BAlightTemp)b,c), and neither was there a clear
pattern across models. Regarding the soil moisture gradient, a group of
models featured decreasing recruitment with increasing soil moisture
(4C, ForClim 1, ForClim11, SIBYLA, xComp, PICUS, LandClim, TreeMig and
iLand), while a few models showed the opposite trend (ForCEEPS(f),
FORMIND and LPJ-GUESS). Across the temperature gradient, all the models
that featured a decrease in recruitment levels with increasing soil
moisture showed the opposite trend (i.e., more recruitment with
increasing temperature), except for 4C. The other models did not feature
a clear pattern.

@@

The EuFoRIa data [@käber2023] are exceptional, particularly with respect
to the number of records (number of sites and repeated measurements on
unmanaged forests), which is essential for capturing a highly 'noisy'
process such as tree regeneration. The use of this dataset for model
benchmarking provided novel insights on the ability of state-of-the-art
models to accurately simulate regeneration levels, species composition,
and mortality in an early stage of tree life, i.e., between tree
diameters of 7 and 10 cm. Overall, by adopting this approach a much
broader understanding resulted than if we had used these data for model
calibration: it is primarily from the shortcomings of the models that we
can gain ecological insights [cf. @trugman2022].

@@

Regeneration levels define the structure and composition of future
forests, but it is equally important to correctly identify areas where
regeneration is lacking [@rammer2021]. There are multiple constraints to
the regeneration niche of tree species [@price2001], and therefore the
absence of regeneration is likely to be common [@fortin2007], even over
larger areas such as the one-hectare plots used here. Tree regeneration
data are characterized by zero-inflation. This was clearly evident from
the EuFoRIa dataset, but some of the models did not produce zeros at
all, or featured a very low proportion of zero data. This substantial
difference may be due to the fact that the simulation results were drawn
from equilibrium forests, whereas in reality many of the forest reserves
are recovering from past management activities and have become denser
over the past decades [e.g., @heiri2009], leading to less regeneration
than in an equilibrium situation. Another possible reason for this
difference may be the exclusion of factors like deer browsing or failure
to accurately measure establishment filters, such as competition with
herb layers, site or environmental limitants.

@@

The species diversity of the entire stand was captured better by the
models than the diversity of tree recruitment, and only a few models
overpredicted stand-level diversity. Defining regeneration composition
as being proportional to adult abundance, regardless of productivity,
might be a simple and conservative assumption to safely maintain
relative species abundances [@hanbury-brown2022], although this may be
simplistic. Furthermore, based on the simulation results, there is no
evidence that models with feedback from the canopy (in terms of species
composition of regeneration) better captured the species diversity of
the recruitment than those without feedback, with the exception of
ForCEEPS that significantly improved regeneration species diversity with
the model version that included feedback (ForCEEPS(f)). The similar
performance of models with and without feedback is likely because the
models put more weight on the regeneration niche arising from abiotic
and biotic filters, than from the habitat niche of the adult trees
[@grubb1977].

@@

There were pronounced differences in how the main tree species were
represented by the models along the environmental gradients, in
particular the dominant species *Abies alba*, *Fagus sylvatica* and
*Picea abies*. The regeneration levels were sampled from simulations in
the equilibrium, and in this state it is expected that non-climax
species such as *Pinus sylvestris* or *Quercus* would be of minor
importance, or absent [@klopcic2015]. Most models captured this low
abundance, which is also found in the empirical data. Thus, although the
broad patterns are matched by many models, improvements in the
quantification of the regeneration niche of the species are needed, but
this cannot be done in the absence of robust datasets across multiple
species.

@@

However, most models properly capture the diversity of the initial tree
community, and differences between model formulations, e.g., the
presence or absence of feedback from the adult trees did not have a
strong effect on capturing the species composition of regeneration.

@@

**Table 1.** Models included in this study, their characteristics and
simulation strategies used. *Scale* refers to the models scale, *Type*
to the models type, *Population structure* to the model population
approach, *Feedback* to the inclusion or not of adult tree feedback to
the level and species composition of regeneration, *Approach* to the
model regeneration module approach, *Start from* to the stage from which
trees are recruited, *Complexity*, to the mean regeneration formulation
as estimated in (Bugmann & Seidl, 2022), *Species* to which species from
the original eleven species were simulated, *Runtime* (spin up) [years]
to the number of years used per model in their simulations for the spin
up period and regeneration sampling; *Climate data type (use of years)*
to the type of data used in the simulations and how it was used, and
*Reference* provides the link to the publication with more model
details. Further information can be found on the individual model
reports available in the supplementary materials.

@@

**Figure 7.** Mean recruitment across the 200 samples per site with the
y-axis scaled differently by model, for the 200 sites against gradients
of (a) total basal area; (b) climatic water balance; (c) seasonal
degree-day sum. The values were split into ten bins; the red lines
represent a GAM model showing the trend in the observed data. Y-axis are
not equally scaled.
